What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-31 android-stacks.tex
discussions and typos

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (141 appplications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and ewas mentioned in approximatelly 
50% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistenly reveal   
fragilities of exception handling code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumente checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robusness not only 
in Android environemt but also to other systems based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications being developed and some of them being largely used. 
Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedent increase of global smartphone shipments -
the mobile industry shiped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

%the mobile industry shiped over 315 million smartphones in the last quarter of 2013
%The pervasiveness of the World Wide Web and the astonishing increase 
%in the number of smartphones stimulated the creation of applications which extend the phones
% capabilities far beyond of the basic calls and textual messages. 
%f we look at global smartphone shipments the numbers are astoning;
%the mobile industry shiped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
%Such increase on smartphones combined with the pervasiveness of the World Wide Web
%have stimulated the creation of applications which extend the phones capabilities 
%far beyond of the basic calls and textual messages.

In this context, Android has become a leading plaftorm for such applictions - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
 own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
and the number of apps built on top of Android increases in a daily rate.

@@

%If on the one hand such applications are getting mode and more popular....
%On the other hand... there is a need for realiability, as some core functionalities 
%are moving to apps....such applications impose more challenges to software reliability 
%Such mobile platforms are intrinsically multi-threaded and support a development model 
% slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],  
%abnormal computation states

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face new threats to application robustness imposed by mobile environmet.
Such applications have to cope with an increasing number of abnormal
computation states, arising from: programmming errors;
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries, compatibility issues [ref], 
conection with devices e.g., as GPS, camera, memory and batery restrictions
among others.

@@

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embeeded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence that the exception handling code may become ilself a source of faults, 
and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.

% Google estimates that Android has over 1 billion active users per month, 
% who pull out theirs phones and check it over 100 bilion times.

@@

Our conjecture is that several Android applicatons may be suffering from similar 
fragilities on the exception handling code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem
 are one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.

@@

Untill recent years, the availability of issues reporting abnormal
computation states (e.g., exception stack traces) was scarce. The
available data comprised only few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their control version as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
on the way exceptions on JNI interface funcions, which can lead to uncaught exceptions that are difficult do 
debug.

@@

Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
Although the study focused on Android opensource applications, some of the results are not restricted to this context, 
for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android plaftorm.
The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
such fragilities in Android and Java environments.

@@

The ramaining of this paper is organized as follows. Section 2 presents a
background on Android Plaform and Java exception model. 
Section 3 presents the study design. Section 4 reports study findings. 
Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 concludes the paper and outlines
directions for future work.

@@

Android is an open source platform for mobile devices based on Linux kernel,
which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
the Dalvik runtime environment, and the Application Framework (which provides 
higher-level services to the applications running on the platform) ~cite{andguide}.
Android platform is mostly implemented in Java and uses JNI[ref] to access native C/C++ code. 
And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
abnormal computation states.

@@

The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent developement (e.g. AsyncTask). 
Such APIs particularly useful due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsivedue due to resurce limitation and excessive network access).

%\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
%  \caption{Android Archtecture} \label{fig:exchier} \end{figure}

@@

%In Java an exception can be thrown in one of the following
%circumstances~\cite{gosling2000java}: 
%(i) explicitly thrown when a throw statement is executed; 
%(ii) implicitly thrown by the JVM when the evaluation of an expression
% violates the normal semantics of language, also refered as a coding error
% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
%OutofMemoryError). 
%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}

@@

% \subsection{Public Project Repositories and Issue Trackers} 
% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
% As mentioned before, currently more and more projects are becomming available in 
% public project repositories such as GitHub and Google Code.
% Such repositories provides facilities such as....
% The issue trackers available in such projects ....

@@

%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
%In Java when an application crashes dut to an uncaught exception ... 
%Exception Stach trace is usually (procurar por artigo).

@@

%\subsection{Research Questions}
In the context of our study we formulated the following reseach questions:

@@

  \item[RQ2] \noindent\emph{Which patterns of exception (mis)-use emerge?}
Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
can pinpoint best practices violations. Chraracteristics of exception stack traces and detected violations can reveal fragilities
 on the exception handling code, which threats the application dependability.

@@

To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
GitHub and Googlecode; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
 reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
(e.g., the application, a library, the Android platform). Next sections describes how such information 
was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
 and how the type associated to each exception and its origens was identified.

@@

%As mentioned before, this study amined at mining information available on GitHub issues,
%more specifically, extracting and distilling the stack traces embedded on GItHub issues.

@@

This set of 482 projects contained overall 32,582 issues from which 4,208 stacktraces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created.

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exceptio related code -  
we opted for not restricting the analysis only on defect
issues.\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}

@@

Different from GitHub, Googlecode does not provide an API to access the information related
 to hosted projects \footnote{Googlecode previoulsly provided a webservice but was deacrivated in June 2013.}.
However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their users to create issues on Googlecode as a way of not loosing the issue history.}.

To overcome this limitation and mine the information available on such Googlecode 
projects we implemented a Web crawler which navegated (incorporated in Exception Miner tool described next) 
 on each project and issue page, extracting the content of each issue to store on relational database. \footnote{Google code previoulsly provided a webservice to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}

To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navegated 
through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

To support this study we developed ExceptionMiner, a modular mining tool able 
to connect to various repositories (such as Google Code,
GHTorrent or even directly to Bugzilla), extract issues, mine stack traces from
them and distill exteption stack trace information. The main components of
ExceptionMiner are as follows:

\noindent\emph{Repository Connectors} The first step in the process is
mining references to exceptions and stack traces embedded in issues, d
GitHub Commector, and GoogleCode Issue Crawler.

@@

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierachy of all exceptions thrown on:
 the last version of SDK version 7, and Android Platform and OS (level 19).

\noindent\emph{Exception Signaler Classifier} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform and OS (level 19)
and this module extractes the packages hierarchies (recursive list of 
of Java packages), that composes it. To discover the packages that compose 
each application, this module extracted the manifest files of each Android apps
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~ref{signalers} summarizes the signaler classification.

@@

%We then analyzed the lables associated to such issues, and we discovered that only 5,196 were
% labled as defect ~\footnote{the dedect lables considered in this work were every lable containing 
%as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the
% issues on which a stack trace was found), most of the issues on which stack traces were found 
%contained no lable. Such observation supports our assumption to consider not only the issues 
% related with defect issues on this analysis.

\subsubsection{Manual Inspections}
This study was also based in a set of manual inspection tasks. Besides the 
 identification of Android projects detailes previously. 
Manual inspections were used discover the type of exceptions thrown by 
the applications and libraries we manually inspect the sourcecode or javadoc
 documentation of each of them. When the exception was removed from the 
last app version, the previous versions were analyzed. When the exception was
 not found on the previous versions, the exception was discarded.  Only 3 exceptions 
remained undefined and were removed from this study.

@@

This section presents the results amed at answering the research 
questions presented before.

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study,
 ranked by the number of distinct project repositories on which they were reported. 
Thes table shows: the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
It also presents the whether the method responsible for signaling the exceptions were
part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
 \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.

@@

We can observe that most of this set is composed by exceptions defined by java.lang,
which are mostly not expliciltly thrown by the developer but include
the exceptions implicitly thrown by the JVM due to programming errors 
(e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError).
For such exceptions, which represent programming bugs or resource limitations, 
there is usually no proper handling besides presenting an error message to
 the user and restarting the application. Only high fault tolerant systems need to 
provide solutions to handle them. Hence, the best strategy to deal with them
is try to prevent them.

@@

 The NullPoiterExceptions are mainly signaled inside Android platform and inside 
application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

@@

Throwing general exceptions, such as direct instances of RuntimeException, violates best
practice III. It is considered a bad practice because exception type does not carry enough
 information to identify the cause of the exceptional behaviour, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

To get a broather view of the root causes of stack traces, we performed an empirical evaluation
whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
to approximately 95\% of all exception types reported on the exception stack-issues.

Such categorization was based known sources of faults mobile developement: concurrency ~cite{to_include}
, backward compatibility ~cite{to_include}, security  ~cite{to_include} and resource management (IO, Memory, Batery) ~cite{to_include}.
We added to this set categories to represent the exceptions caused by programming mistakes,
 and general Java exceptions (whose types not convey any specific cause of failure), and 
exceptions that represent specific  exceptional situations of apps, frameworks or libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

@@

In this second analysis the programming mistakes were also the
most mentioned root causes, followed by exceptions related to resource 
management (which was expected since on mobile environment 
resources are usualy scarse). We could observe that there were 
stack traces related to the main concerns of mobile develoment.

@@

\subsection{What the Types of Root Exceptions can tell us?}
%\subsection{Exploring the Information Embeeded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

%\begin{figure}
%\centering
%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
%\includegraphics[width=\hsize]{exception_types.pdf}
%\caption{Distribution of Checked and Unchedked exceptions.}
%\label{fig:typeroot}
%\end{figure}

@@

In order to evaluate if  best practice IV was being obeyed, we manualy inspected the   
signature of each method signaling a runtime exception, that was defined by third 
party code (lib or Android) and reused by an application.
In this task we filtered out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) since these exceptions should not be documented on the 
method signature.

@@

We could also observe checked exceptions (which according to Java specification represent
recoverable conditions) as the root causes of reported exception stack traces.
A checked exception can remain uncaught if it is declared by evey method on the call chain or if 
it is wrapped in a runtime exception and rethrown. Next, section describes the kinds of 
wrappings found in this study.

@@

\subsection{Exploring the Information Embeeded on Exception Wrappings}
Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors.

@@

Taking a closer look on the methods reposponsible for such wrappings,
 we could observe that aproximately 41% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.

@@

Figure X presents a code snippet of ActivityThreat which is reponsible for (among other things)
calling the Activities onCreate method. Hence, as we can see from lines 4-8,
 in case any exception is thrown by an application Activity during its creation, 
it is wrapped in a general Runtime exception and rethrown.

>>>>

This paper reports a repository mining study which analyzed all the issues created on
Android projects available in GitHub (482 applications) and Google code (141 applications)
Overall, our research set includes XXXX issues from which XXX stack traces 
were extracted. Such stack traces were distilled and analyzed in 
combination with source code and bytecode analysis. This study 
consistently identifies common characteristics of reported 
exception stacks traces - e.g., the NullPointerException was the reported cause 
on 30,08\% of all analyzed exception stack traces, and was mentioned in approximately 
50% of all projects analyzed. Moreover,  the cross-project analysis 
of exception stack traces enabled to consistently reveal   
fragilities of exception handling code: 
(1) undocumented runtime exceptions signaled by third party code,
(2) undocumented checked exceptions thrown by Android native code
(3) unexpected exception wrappings (e.g., Errors being wrapped in checked exceptions) 
revealing that Java hybrid exception model is not fully used according to its purpose.
Such fragilities impose challenges to application robustness not only 
in Android environment but also to other systems based on Java exception model.

@@

Along the recent years, we have witnessed an astonishing increase in the number 
mobile applications being developed and some of them being largely used. 
Such increase is motivated by pervasiveness of the World Wide Web 
and the unprecedented increase of global smartphone shipments -
the mobile industry shipped over 315 million smartphones in the last quarter of 2013 ~\cite{googleio}.

%the mobile industry shipped over 315 million smartphones in the last quarter of 2013
%The pervasiveness of the World Wide Web and the astonishing increase 
%in the number of smartphones stimulated the creation of applications which extend the phones
% capabilities far beyond of the basic calls and textual messages. 
%f we look at global smartphone shipments the numbers are astonishng;
%the mobile industry shipped over 315 million smartphones in the last quarter of 2013 [ref-googleIO].
%Such increase on smartphones combined with the pervasiveness of the World Wide Web
%have stimulated the creation of applications which extend the phones capabilities 
%far beyond of the basic calls and textual messages.

In this context, Android has become a leading platform for such applications - thanks to its open-source 
model and hardware partners like Samsung, HTC, Motorola, Asus (besides Nexus Google's
 own device) ~\cite{gartner}.  Google estimates that there is over 1 billion Android active users per month ~\cite{googleio},
and the number of apps built on top of Android increases in a daily rate.

@@

%If on the one hand such applications are getting mode and more popular....
%On the other hand... there is a need for reliability, as some core functionalities 
%are moving to apps....such applications impose more challenges to software reliability 
%Such mobile platforms are intrinsically multi-threaded and support a development model 
% slightly (or sometimes drastically) different from those adopted on Web development [arie-icse],  
%abnormal computation states

In the same rate the number of a users increase, also increases 
the number of users affected by application faults and crashes.
If on the one hand such applications extend phones capabilities 
far beyond of the basic calls and textual messages, on the other hand
they have to face new threats to application robustness imposed by mobile environment.
Such applications have to cope with an increasing number of abnormal
computation states, arising from: programming errors;
 noisy user inputs; faults in underlying middleware or hardware; 
faults on third party services and libraries, compatibility issues [ref], 
connection with devices e.g., as GPS, camera, memory and battery restrictions
among others.

@@

Therefore, techniques for error detection and handling are not  an optional add-on but a 
fundamental part of such apps.
% as any other modern system~\cite{bruntink2006discovering}. 
The Android platform uses the exception handling mechanism embedded
 in Java to deal with such conditions. Although the exception handling mechanism
~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions, studies (e.g.~\cite{miller1997issues,Robil00,shah2010understanding, garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
have shown that it is generally poorly understood and the less tested parts of the system ~\cite{coelho2011unveiling}.
As a consequence that the exception handling code may become itself a source of faults, 
and may affect the system the other way around, making the system even less robust ~\cite{robillard2003static,coelho2011unveiling}.

% Google estimates that Android has over 1 billion active users per month, 
% who pull out theirs phones and check it over 100 billion times.

@@

Our conjecture is that several Android applications may be suffering from similar 
fragilities on the exception handling code (of the applications themselves or 
the underlaying platform), and that such fragilities may be leading to
decrease the application robustness, leading for instance to the Uncaught Exception problem
 are one of the main causes of application crashes ~\cite{jo2004uncaught} ~\cite{zhang2012amplifying}.

@@

Until recent years, the availability of issues reporting abnormal
computation states (e.g., exception stack traces) was scarce. The
available data comprised only few large scale projects (e.g. the Mozilla
crash report dataset\footnote{https://code.google.com/p/promisedata/issues/detail?id=88}).
Recently we have observed a rapid increase in the number of open
source projects that make their control version as well as the issue tracking systems
publicly available in repository hosting sites such as GitHub and Googlecode.

@@

   \item Undocumented checked exception signaled by native Android library. This finding points to a fragility 
on the way exceptions on JNI interface functions, which can lead to uncaught exceptions that are difficult do 
debug.

@@

Such fragilities represent obstacles to the development of dependable java-based applications as they make difficult and 
sometimes almost impossible to protect the code against the Uncaught Exceptions that lead to system crashes.
Although the study focused on Android open-source applications, some of the results are not restricted to this context, 
for instance: some of the unexpected wrappings were performed by Java libraries and elements (e.g. reflection library and 
the static blocks); and the undocumented runtime exceptions were also found in Java libraries besides the Android platform.
The study results are relevant to: (i) Android and Java developers to familiarize common characteristics and fragilities of exception-related code 
which is the first step to help developers preventing and dealing with them; (ii) designers of languages and tools to consider ways of reducing 
such fragilities in Android and Java environments.

@@

The remaining of this paper is organized as follows. Section 2 presents a
background on Android Platform and Java exception model. 
Section 3 presents the study design. Section 4 reports study findings. 
Section 5 provides further discussions and insights.
Section 6 presents the threats to validity associated to this study. Finally Section
7 describes the related work, and Section 8 concludes the paper and outlines
directions for future work.

@@

Android is an open source platform for mobile devices based on Linux kernel,
which also contains  a core Java library, a set of libraries third-party libraries (e.g.  SQLite),
the Dalvik runtime environment, and the Application Framework (which provides 
higher-level services to the applications running on the platform) ~\cite{andguide}.
Android platform is mostly implemented in Java and uses JNI[ref] to access native C/C++ code. 
And as mentioned before, it reuses the embedded Java Exception model to detect and handle 
abnormal computation states.

@@

The Android Application Framework also provides several supporting APIs, such as 
the APIs to support concurrent development (e.g. AsyncTask). 
Such APIs particularly useful due to the high demand for responsiveness 
in mobile environment (mobile apps can easily become 
unresponsive due to resource limitation and excessive network access).

%\begin{figure} \centering \includegraphics[scale=0.8]{arch3.png}
%  \caption{Android Architecture} \label{fig:exchier} \end{figure}

@@

%In Java an exception can be thrown in one of the following
%circumstances~\cite{gosling2000java}: 
%(i) explicitly thrown when a throw statement is executed; 
%(ii) implicitly thrown by the JVM when the evaluation of an expression
% violates the normal semantics of language, also referred as a coding error
% (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
%(iii) implicitly thrown by the JVM due to an internal error or resource limitation (e.g.,
%OutofMemoryError). 
%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}

@@

% \subsection{Public Project Repositories and Issue Trackers} 
% ONE OF THE REVIEWERS COMPLAINED ABOUT THE ABSENCE OF EXPLANATIONS OF WHAT AN ISSU IS...
% As mentioned before, currently more and more projects are becoming available in 
% public project repositories such as GitHub and Google Code.
% Such repositories provides facilities such as....
% The issue trackers available in such projects ....

@@

%This study focus on the analysis of crash information (related to uncaught exceptions) of a family of Android applications
%aiming at  extracting their common characteristics and looking for similar vulnerabilities on applications and underlaying platform.
%In Java when an application crashes due to an uncaught exception ... 
%Exception Stach trace is usually (procurar por artigo).

@@

%\subsection{Research Questions}
In the context of our study we formulated the following research questions:

@@

  \item[RQ2] \noindent\emph{Which patterns of exception (mis)-use emerge?}
Based on a set of best practices compiled in this study, the study investigates whether characteristics on exception stack traces  
can pinpoint best practices violations. Characteristics of exception stack traces and detected violations can reveal fragilities
 on the exception handling code, which threats the application dependability.

@@

To answer the research questions the following information is needed: (i) the issues of reported on every Android project hosted on 
GitHub and Googlecode; (ii) the stack traces embedded on those issues; (iii) the type of the exceptions
 reported on the stack traces (e.g., error, runtime, checked); (iv) the origin of such exceptions 
(e.g., the application, a library, the Android platform). Next sections describes how such information 
was retrieved from GitHub and Googlecode, how the exceptions stack traces were extracted and distilled,
 and how the type associated to each exception and its origins was identified.

@@

%As mentioned before, this study mined information available on GitHub issues,
%more specifically, extracting and distilling the stack traces embedded on GItHub issues.

@@

This set of 482 projects contained overall 32,582 issues from which 4,208 exception stack traces 
were extracted. Issues on Github are different from issues on dedicated bug tracking tools such as 
Bugzilla and Jira. The most important difference is that there are no predefined fields
  (e.g. severity and priority). Instead, Github uses a more open ended tagging system, on which
repositories are offered a pre-defined set of labels, but repository owners can modify 
them at will. Therefore, an issue may have none ore an arbitrary set of labels depending 
on which repository it was created.

Based on the assumption that regardless the issue type, every exception stack
trace contains relevant information concerning the exception structure of the
projects analyzed - and therefore can review fragilities on the exception-related code -  
we opted for not restricting the analysis only on defect
issues.\footnote{We conducted the same analysis on the defect issues and the top
exceptions are similar to the ones mentioned on defect issues. Due to space
limitation we limit to present the general analysis here. More detailed analysis
on the defect issues can be found at:
\url{www.dimap.ufrn.br/~roberta/android_repo_mining}}

@@

Different from GitHub, Googlecode does not provide an API to access the information related
 to hosted projects \footnote{Googlecode previoulsly provided a Web service but was deactivated in June 2013.}.
However many well used open source Android application such as K9mail, is hosted in Googlecode which stores
the issue history of this and other apps are store there ~\footnote{K9mail moved to Github but advice their 
users to create issues on Googlecode as a way of not loosing the issue history.}.

To overcome this limitation and mine the information available on such Googlecode 
projects we implemented a Web crawler which navigated (incorporated in Exception Miner tool described next) 
 on each project and issue page, extracting the content of each issue to store on relational database. \footnote{Google code previously provided a Web service to extract information about each project, but this service was removed in what Google called "clean-up action" which happened in June 2013.}

To overcome this limitation we needed to implement a Web Crawler (incorporated in Exception Miner tool described next) that navigated 
through the web interface of Googlecode projects extracting all issues and issue comments and storing in a relational database for later analysis.

@@

To support this study we developed ExceptionMiner, a modular mining tool able 
to connect to repositories, extract issues, mine stack traces from
them and distill exception stack trace information. The main components of
ExceptionMiner are as follows:

\noindent\emph{Repository Connectors}  This component is responsible to connect to
 the host repositories in order to collect project issues. The GitHub Connector connected
 to GiHub through GHTorrent. The GoogleCode connector comprised of a Web Crawler,
 that traversed Googlecode Web interface in order to extract the project issues.

@@

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierarchy of all exceptions thrown on:
 the last version of SDK version 7, and Android Platform and OS (level 19).

\noindent\emph{Exception Signaler Classifier} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform and OS (level 19)
and this module extracts the packages hierarchies (recursive list of 
of Java packages), that composes it. To discover the packages that compose 
each application, this module extracted the manifest files of each Android apps
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~ref{signalers} summarizes the signaler classification.

@@

%We then analyzed the labels associated to such issues, and we discovered that only 5,196 were
% labled as defect ~\footnote{the dedect lables considered in this work were every lable containing 
%as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the
% issues on which a stack trace was found), most of the issues on which stack traces were found 
%contained no lable. Such observation supports our assumption to consider not only the issues 
% related with defect issues on this analysis.

\subsubsection{Manual Inspections}
This study was also based in a set of manual inspection tasks. Besides the 
 identification of Android projects details previously. 
Manual inspections were used discover the type of exceptions thrown by 
the applications and libraries we manually inspect the sourcecode or javadoc
 documentation of each of them. When the exception was removed from the 
last app version, the previous versions were analyzed. When the exception was
 not found on the previous versions, the exception was discarded.  Only 3 exceptions 
remained undefined and were removed from this study.

@@

This section presents the results aimed at answering the research 
questions presented before.

@@

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:toptenandroid} illustrates the top 10 root causes found in the study,
 ranked by the number of distinct project repositories on which they were reported. 
The table shows: the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectively.
It also presents the whether the method responsible for signaling the exceptions were
part of Android, the application itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
 \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.

@@

We can observe that most of this set is composed by exceptions defined by java.lang,
which are mostly not explicitly thrown by the developer but include
the exceptions implicitly thrown by the JVM due to programming errors 
(e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError).
For such exceptions, which represent programming bugs or resource limitations, 
there is usually no proper handling besides presenting an error message to
 the user and restarting the application. Only high fault tolerant systems need to 
provide solutions to handle them. Hence, the best strategy to deal with them
is try to prevent them.

@@

 The NullPointerExceptions are mainly signaled inside Android platform and inside 
application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

@@

Throwing general exceptions, such as direct instances of RuntimeException, violates best
practice III. It is considered a bad practice because exception type does not carry enough
 information to identify the cause of the exceptional behavior, and as a consequence developers need to relying on
the exception message which can be neither complete nor precise~\cite{gosling2000java}.

@@

To get a broader view of the root causes of stack traces, we performed an empirical evaluation
whose goal was to categorized the top 100 root exceptions. The top 100 root causes correspond 
to approximately 95\% of all exception types reported on the exception stack-issues.

Such categorization was based known sources of faults mobile development: concurrency [INCLUDEREF]
, backward compatibility  [INCLUDEREF], security   [INCLUDEREF] and resource management (IO, Memory, Batery)  [INCLUDEREF].
We added to this set categories to represent the exceptions caused by programming mistakes,
 and general Java exceptions (whose types not convey any specific cause of failure), and 
exceptions that represent specific  exceptional situations of apps, frameworks or libraries.
The list of categories is illustrated in Table~\ref{tab:categories}.

@@

In this second analysis the programming mistakes were also the
most mentioned root causes, followed by exceptions related to resource 
management (which was expected since on mobile environment 
resources are usually scarce). We could observe that there were 
stack traces related to the main concerns of mobile develoment.

@@

\subsection{What the Types of Root Exceptions can tell us?}
%\subsection{Exploring the Information Embedded on Exception Types}
%\subsection{The Exception Types and What They can Tell us? }

@@

%\begin{figure}
%\centering
%\includegraphics[width=\hsize]{chart_exceptiontypes.pdf}
%\includegraphics[width=\hsize]{exception_types.pdf}
%\caption{Distribution of Checked and Unchecked exceptions.}
%\label{fig:typeroot}
%\end{figure}

@@

In order to evaluate if  best practice IV was being obeyed, we manually inspected the   
signature of each method signaling a runtime exception, that was defined by third 
party code (lib or Android) and reused by an application.
In this task we filtered out all the exceptions implicitly 
signaled by  JVM (due to programming mistakes) since these exceptions should not be documented on the 
method signature.

@@

We could also observe checked exceptions (which according to Java specification represent
recoverable conditions) as the root causes of reported exception stack traces.
A checked exception can remain uncaught if it is declared by every method on the call chain or if 
it is wrapped in a runtime exception and re-thrown. Next, section describes the kinds of 
wrappings found in this study.

@@

\subsection{Exploring the Information Embedded on Exception Wrappings}
Table~\ref{tab:wrappingandroid} presents the wrappings found in this study for all
stack traces found in Android repositories. As we can see, some of the checked
exceptions were indeed wrapped in runtime exceptions or even errors.

@@

Taking a closer look on the methods responsible for such wrappings,
 we could observe that approximately 41% of them were performed by a single class,
 named ActivityThreat. This class is responsible for managing the life cycle of an Android Activities.

@@

Figure X presents a code snippet of ActivityThreat which is responsible for (among other things)
calling the Activities onCreate method. Hence, as we can see from lines 4-8,
 in case any exception is thrown by an application Activity during its creation, 
it is wrapped in a general Runtime exception and re-thrown.
